### 2022-01-07 LaTeX/tarjan.tex
Slides updates, figures added
Updates

\hypersetup{
    colorlinks=true,
    linktoc=false,
    linkcolor=blue
}

@@

\subsubsection{Backtracking edges}
\begin{definition}
    Given a graph \GG and an order of traversal in this graph, \textit{i.e.} each node of \GG has a unique value $\texttt{num} \in [0,~|\mathcal{V}|]~$ and two nodes $u$ and $v$, there is a backtracking edge from $v$ to $u$ if:
    \begin{equation*}
        \left\{
            \begin{array}{l}
                \texttt{u.num} < \texttt{v.num}\\
                v~\leadsto~u,~\textit{i.e. }~(v, u) \in \mathcal{E}\\
            \end{array}
        \right.
    \end{equation*}
    In this case, the backtracking edge from $v$ to $u$ is represented by $v \hookrightarrow u$.\\      
\end{definition}

@@

\pagebreak
\section{Tarjan's algorithm}
Tarjan's algorithm is an efficient on-the-fly SCC computing algorithm. It basically perfoms a DFS while updating the \textit{num} and \textit{lowlink} values. All nodes are stored in a stack during the traversal until a backtracking edge is found. In this case, the \textit{lowlinks} are computed and all nodes are unstacked and saved in a SCC until a node verifying the equality between its \textit{num} and its \textit{lowlink} -- which has to occur -- is found. Then, the DFS goes on. The whole process is written in the following algorithm \ref{alg:tarjan}.\\
It can be shown that every node and edge are visited only once so the algorithm can achieve a linear complexity, \textit{i.e.} $\mathcal{O}(|\mathcal{V}|+|\mathcal{E}|)$.

@@

\subsection{The algorithm}
\begin{algorithm}[H]\label{alg:seqsetbased}
    \SetAlgoLined
    \KwData{A graph \GG = (\VV, \EE), a starting node $v_0$\;}
    \KwResult{A partition \texttt{SCCs} of \VV where each element of \texttt{SCCs} is a maximal set of strongly connected components of \GG\;}
    Initialize an empty set \texttt{DEAD}\;
    Initialize an empty set \texttt{VISITED}\;
    Initialize an empty stack \texttt{R}\;
    setBased($v_0$)\;
    \SetKwProg{Function}{function}{}{}
    \Function{setBased: $v \in \mathcal{V} \rightarrow \texttt{None}$}{
        $\texttt{VISITED} := \texttt{VISITED} \cup \{v\}$\;
        \texttt{R.push($v$)}\;
        \ForEach{$w\in \texttt{POST(v)}$}{
            \If{$w\in \texttt{DEAD}$}{
                continue\;
            }
            \ElseIf{$w \notin \texttt{VISITED}$}{
                setBased($w$)\;
            }
            \Else{
                \While{$\mathcal{S}(v) \neq \mathcal{S}(w)$}{
                    $r := \texttt{R.pop()}$\;
                    $\texttt{UNITE}(\mathcal{S}, r, \texttt{R.top()})$\;
                }
            }
        }
        \If{$v = \texttt{R.top()}$}{
            \textbf{report SCC} $\mathcal{S}(v)$\;
            $\texttt{DEAD} := \texttt{DEAD} \cup \mathcal{S}(v)$\;
            \texttt{R.pop()}\;
        }
    }

@@

\subsection{Isabelle proof}
cf Proof

%-----------------------------------------------------------------
\pagebreak

@@

\bibitem{} R. Chen, C. Cohen, J.-J. Lévy, S. Merz, L. Théry, \emph{Formal Proofs of Tarjan’s Strongly Connected 2 Components Algorithm in Why3, Coq and Isabelle}, 2019

\bibitem{} V. Bloemen, A. Laarman, J. van de Pol, \emph{Multi-Core On-The-Fly SCC Decomposition}, 2016

\bibitem{} V. Bloemen, \emph{Strong Connectivity and Shortest Paths for Checking Models}, 2019

>>>>

\hypersetup{
    colorlinks=true,
    linktoc=false,
    linkcolor=blue,
    citecolor=blue
}

@@

\subsubsection{Backtracking edges}
\begin{definition}
    Given a graph \GG and an order of traversal in this graph, \textit{i.e.} each node of \GG has a unique value $\texttt{num} \in [0,~|\mathcal{V}|]~$ and two nodes $u$ and $v$, there is a backtracking edge from $v$ to $u$ if:
    \begin{equation*}
        \left\{
            \begin{array}{l}
                \texttt{u.num} < \texttt{v.num}\\
                v~\Rightarrow~u,~\textit{i.e. }~(v, u) \in \mathcal{E}\\
            \end{array}
        \right.
    \end{equation*}
    In this case, the backtracking edge from $v$ to $u$ is represented by $v \hookrightarrow u$.\\      
\end{definition}

@@

\pagebreak
\section{Tarjan's algorithm}
Tarjan's algorithm is an efficient on-the-fly SCC computing algorithm \cite{tarjanMerz}. It basically perfoms a DFS while updating the \textit{num} and \textit{lowlink} values. All nodes are stored in a stack during the traversal until a backtracking edge is found. In this case, the \textit{lowlinks} are computed and all nodes are unstacked and saved in a SCC until a node verifying the equality between its \textit{num} and its \textit{lowlink} -- which has to occur -- is found. Then, the DFS goes on. The whole process is written in the following algorithm \ref{alg:tarjan}.\\
It can be shown that every node and edge are visited only once so the algorithm can achieve a linear complexity, \textit{i.e.} $\mathcal{O}(|\mathcal{V}|+|\mathcal{E}|)$.

@@

\subsection{The algorithm}
See \cite{articleSCC} for the original paper.\\
\begin{algorithm}[H]\label{alg:seqsetbased}
    \SetAlgoLined
    \KwData{A graph \GG = (\VV, \EE), a starting node $v_0$\;}
    \KwResult{A partition \texttt{SCCs} of \VV where each element of \texttt{SCCs} is a maximal set of strongly connected components of \GG\;}
    Initialize an empty set \texttt{DEAD}\;
    Initialize an empty set \texttt{VISITED}\;
    Initialize an empty stack \texttt{R}\;
    setBased($v_0$)\;
    \SetKwProg{Function}{function}{}{}
    \Function{setBased: $v \in \mathcal{V} \rightarrow \texttt{None}$}{
        $\texttt{VISITED} := \texttt{VISITED} \cup \{v\}$\;
        \texttt{R.push($v$)}\;
        \ForEach{$w\in \texttt{POST(v)}$}{
            \If{$w\in \texttt{DEAD}$}{
                continue\;
            }
            \ElseIf{$w \notin \texttt{VISITED}$}{
                setBased($w$)\;
            }
            \Else{
                \While{$\mathcal{S}(v) \neq \mathcal{S}(w)$}{
                    $r := \texttt{R.pop()}$\;
                    $\texttt{UNITE}(\mathcal{S}, r, \texttt{R.top()})$\;
                }
            }
        }
        \If{$v = \texttt{R.top()}$}{
            \textbf{report SCC} $\mathcal{S}(v)$\;
            $\texttt{DEAD} := \texttt{DEAD} \cup \mathcal{S}(v)$\;
            \texttt{R.pop()}\;
        }
    }

@@

\subsection{Isabelle proof}
Not finished yet! See my oral presentation on January 14, 2022.
%-----------------------------------------------------------------
\pagebreak

@@

\bibitem{tarjanMerz} R. Chen, C. Cohen, J.-J. Lévy, S. Merz, L. Théry, \emph{Formal Proofs of Tarjan’s Strongly Connected 2 Components Algorithm in Why3, Coq and Isabelle}, 2019

\bibitem{thesisBloemen} V. Bloemen, A. Laarman, J. van de Pol, \emph{Multi-Core On-The-Fly SCC Decomposition}, 2016

\bibitem{articleSCC} V. Bloemen, \emph{Strong Connectivity and Shortest Paths for Checking Models}, 2019
